---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file and re-knit with devtools::build_readme(). -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	collapse = TRUE,
	comment = "#>",
	fig.path = "man/figures/README-",
	fig.width = 7,
	fig.height = 4,
	dpi = 150,
	out.width = "100%"
)
options(width = 100)
```

# Fused Extended Two-Way Fixed Effects

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/fetwfe)](https://CRAN.R-project.org/package=fetwfe)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/fetwfe)](https://CRAN.R-project.org/package=fetwfe)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
<!-- badges: end -->

The `{fetwfe}` package implements *fused extended two-way fixed effects* (FETWFE), a methodology for estimating treatment effects in difference-in-differences with staggered adoptions.

In staggered-adoption settings — where units become treated at different times — treatment effects typically vary across adoption cohorts and over time since treatment. The conventional two-way fixed effects (TWFE) estimator is biased when effects are heterogeneous like this, because it effectively uses already-treated units as controls for newly-treated ones, contaminating the estimate.

FETWFE instead fits a *fully heterogeneous* model — a separate effect for every cohort and time period — and applies a **fusion penalty** that shrinks together effects which are similar across neighboring cohorts and periods. This pools the cohort-by-time effects in a data-driven way (collapsing them where they genuinely agree, keeping them distinct where they differ) and, unlike ad hoc model selection, comes with **asymptotically valid standard errors and confidence intervals** for the resulting estimates.

* For a brief introduction to the methodology, as well as background on difference-in-differences with staggered adoptions and motivation for FETWFE, see this [blog post](https://gregoryfaletto.com/2023/12/13/new-paper-fused-extended-two-way-fixed-effects-for-difference-in-differences-with-staggered-adoptions/).
* For more detailed slides on the methodology (but less detailed than the paper), see [here](https://gregoryfaletto.com/2024/02/11/presentation-on-fused-extended-two-way-fixed-effects/).
* Check out the most recent draft of the full paper [here](https://arxiv.org/abs/2312.05985).
* This [blog post](https://gregoryfaletto.com/2025/01/03/new-r-fetwfe-package-implementing-fused-extended-two-way-fixed-effects/) explains a little more about what the package does under the hood, if you're interested.

## Installation

To install the released version of `{fetwfe}` from CRAN:

```r
install.packages("fetwfe")
```

Or the latest development version from GitHub:

```r
# install.packages("remotes")  # if needed
remotes::install_github("gregfaletto/fetwfePackage")
```

## A worked example: fit, extract, plot

The primary function is `fetwfe()`. Here it is applied to the `divorce` dataset from the `bacondecomp` package — the unilateral ("no-fault") divorce / female-suicide panel of Stevenson and Wolfers (2006), reproducing the empirical application in Faletto (2025, Sec. 8.2):

```{r fit, message = FALSE, warning = FALSE}
library(fetwfe)
library(bacondecomp)

data(divorce)
# Female subset; `changed` is the absorbing 0/1 divorce-reform indicator, and the
# response is the elasticity-scaled female suicide rate. Noise variances are
# supplied (precomputed by REML) to keep the call fast.
divorce_f <- divorce[divorce$sex == 2, ]

res <- fetwfe(
	pdata = divorce_f,
	time_var = "year",
	unit_var = "st",
	treatment = "changed",
	covs = c("murderrate", "lnpersinc", "afdcrolls"),
	response = "suiciderate_elast_jag",
	sig_eps_sq = 0.0344,
	sig_eps_c_sq = 0.1507,
	add_ridge = TRUE,
	q = 0.5
)
```

**1. The overall ATT.** The aggregated effect and its 95% Wald interval:

```{r overall}
round(c(
	ATT = res$att_hat,
	SE = res$att_se,
	CI_low = res$att_hat - qnorm(0.975) * res$att_se,
	CI_high = res$att_hat + qnorm(0.975) * res$att_se
), 4)
```

(`summary(res)` prints the full report — overall ATT, per-cohort and event-study effects, and model details.)

**2. The heterogeneity underneath.** The overall ATT is an average; `cohortStudy()` and `eventStudy()` return the effects by *adoption cohort* and by *time since treatment* as tidy data frames — the effects FETWFE kept distinct rather than fusing to zero:

```{r extract}
cohortStudy(res)
```

**3. Plot it.** `plot()` draws the event-study estimates with confidence intervals (or per-cohort average effects with `type = "catt"`):

```{r event-study-plot, fig.alt = "Event-study plot: estimated effect of no-fault divorce reform on the elasticity-scaled female suicide rate, by years since adoption, with confidence intervals."}
plot(res, type = "event_study")
```

For the full set of vignettes and function documentation, see the [package website](https://gregfaletto.github.io/fetwfePackage/) and the [CRAN page](https://CRAN.R-project.org/package=fetwfe).

## Penalty and fusion structures

A core differentiator of FETWFE is the *geometry* of its fusion penalty, chosen via the `fusion_structure` argument of `fetwfe()`:

- **`"cohort"`** (the default) fuses treatment effects within and between cohorts — the two-way fusion structure of Faletto (2025).
- **`"event_study"`** instead fuses effects at the same time since treatment (event time `e = t - g`) across cohorts — the package realization of the paper's event-study-penalty theory.

For full control, the `fusion_matrix` argument accepts any invertible `num_treats x num_treats` differencing matrix `D_N`, encoding an arbitrary fusion structure beyond the two built-ins. The same `fusion_structure` option is available in `genCoefs()` for simulation studies.

For guidance on which to use, see the vignette *"Choosing a fusion structure: cohort vs. event-study penalties"* — `vignette("fusion_structure_vignette", package = "fetwfe")`.

## References
- Faletto, G (2025). *Fused Extended Two-Way Fixed Effects for Difference-in-Differences with Staggered Adoptions*. [arXiv preprint arXiv:2312.05985](https://arxiv.org/abs/2312.05985).
