---
title: "Choosing a fusion structure: cohort vs. event-study penalties"
author: "Gregory Faletto"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    math_method: "mathml"
    output_file: "FETWFE_Fusion_Structure_Vignette.html"
vignette: >
  %\VignetteIndexEntry{Choosing a fusion structure: cohort vs. event-study penalties}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2, dplyr}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(fetwfe)
```

# What a "fusion structure" is

FETWFE estimates a separate treatment effect for every treated cohort `g` in
every post-treatment period `t` --- written `tau_{g, t}` --- and then shrinks
those effects toward each other with a *fusion penalty*. "Fusion" means the
penalty acts on the *differences* between selected pairs of treatment-effect
coefficients: pushing a difference toward zero pools the two effects, so the
estimator borrows strength across cells instead of estimating each `tau_{g, t}`
in isolation. Which differences the penalty acts on is the *fusion structure*,
and it encodes the kind of similarity you expect the treatment effects to have.

Since version 1.19.0 `fetwfe()` exposes this choice through a single argument,
`fusion_structure`, with two options:

```{r, eval = FALSE}
fetwfe(..., fusion_structure = "cohort")       # the default
fetwfe(..., fusion_structure = "event_study")  # the alternative added for issue #40
```

Both penalties are *data-driven*, not hard restrictions. The estimator pools
toward the chosen structure but is free to deviate from it when the data demand
it. So choosing a fusion structure does **not** compromise the asymptotic
validity of the estimator (a lemma in the paper's extensions section shows the
event-study penalty satisfies the same conditions, with the same singular-value
constants `sigma_max <= sqrt(6)` and `sigma_min >= 1 / (T * sqrt(2 * T))`, as the
default; every consistency, restriction-selection, and inference guarantee in the
paper therefore holds verbatim for it). The cost of guessing "wrong" is
*finite-sample efficiency* --- noisier estimates --- not bias or invalid
confidence intervals.

## `fusion_structure = "cohort"` (the default)

The default penalty fuses **adjacent cohorts at the same calendar time** and
**adjacent time periods within the same cohort**. It pools toward the assumption
that treatment effects vary *smoothly* across neighboring cohorts and smoothly
over calendar time within a cohort. This is a sensible, agnostic default: it does
not privilege any single summary of the effects.

## `fusion_structure = "event_study"`

The event-study penalty instead fuses, for each *event time*
`e = 0, 1, 2, ...` (time since treatment), the effect `tau_{r, r + e}` across all
cohorts `r`. It pools toward the assumption that treatment effects depend only on
**time since treatment**, not on the calendar period or the cohort --- the
identifying restriction behind the standard event-study aggregation used
throughout applied difference-in-differences (e.g.,
`did::aggte(type = "dynamic")`). When the data really are driven by event time
(gradual onset, a dose-response that accumulates with exposure), this pools
strength across cohorts at each event time and estimates the dynamic path more
efficiently.

# Two meanings of "event study" --- read this first

The word "event study" appears in `fetwfe` in **two distinct, independent
places**, and conflating them is the most common source of confusion:

- **`fusion_structure = "event_study"`** is an *estimation-time penalty*. It
  changes the **fit** --- the coefficients you get out of `fetwfe()`.
- **`family = "event_study"`** in `eventStudy()` and
  `simultaneousCIs(fit, family = "event_study")` is a *reporting-time
  aggregation*. It changes how an already-fitted set of effects is
  **summarized** (one number per event time), and does nothing to the fit.

These two choices are **orthogonal**: you can report an event-study aggregation
of a *cohort*-penalized fit, or a cohort aggregation of an *event-study*-penalized
fit. Every combination is valid:

```{r, eval = FALSE}
# event-time SUMMARY of a default (cohort-penalized) fit:
eventStudy(fetwfe(..., fusion_structure = "cohort"))

# cohort SUMMARY of an event-study-penalized fit:
simultaneousCIs(fetwfe(..., fusion_structure = "event_study"), family = "cohort")
```

The rest of this vignette uses `eventStudy()` (the reporting aggregation) to
*display* fits made under both penalties --- a concrete illustration of the
orthogonality.

# A side-by-side demo

We simulate a panel whose treatment effects are **sparse in the event-study
basis** --- effects that share the same time since treatment (event time `e`)
are fused across cohorts, so the true effect depends mostly on `e`. We generate
this truth natively with `genCoefs(fusion_structure = "event_study")` (the
simulation-side companion to the estimator argument), then fit the panel with
both fusion structures and see which recovers the truth better. Because we know
the data-generating effects exactly, we can measure recovery error directly.

We use the package's simulation utilities (`genCoefs()` / `simulateData()` /
`fetwfeWithSimulatedData()`; see the *Simulation* vignette for a full tour). The
treatment-effect coefficients occupy a known block of the coefficient vector;
`fetwfe()` reports their positions in its `treat_inds` slot (see `?fetwfe`), and
those positions depend only on the panel shape (`G` cohorts, `T` periods, `d`
covariates):

```{r}
G <- 6L
T <- 9L
d <- 1L
N <- 400L

# Generate truth that is sparse in the event-study basis directly, with
# fusion_structure = "event_study". seed 10 is a representative draw (see the
# note after the plot); density 0.25 keeps the truth genuinely sparse in the
# event-study basis.
coefs <- genCoefs(
  G = G, T = T, d = d, density = 0.25, eff_size = 2, seed = 10,
  fusion_structure = "event_study"
)

# Locate the treatment-effect block tau_{g, t} (cohort-major order).
cohort_times <- getTes(coefs)$cohort_times # each cohort's first treated period
n_k <- T - cohort_times + 1 # number of post-treatment effects per cohort
num_treats <- sum(n_k)
base_cols <- G + (T - 1) + d + d * G + d * (T - 1) # fixed effects + covariate terms
treat_inds <- base_cols + seq_len(num_treats) # the tau block

# Event time (time since treatment) of each tau, in the same order.
event_time <- unlist(lapply(n_k, function(nk) 0:(nk - 1)))

# The true per-cell treatment effects, generated natively (not hand-built).
true_te <- coefs$beta[treat_inds]

sim <- simulateData(coefs, N = N, sig_eps_sq = 1, sig_eps_c_sq = 0.5)
```

Now fit the same simulated panel under each fusion structure:

```{r, message = FALSE, warning = FALSE}
fit_cohort <- fetwfeWithSimulatedData(sim, lambda_selection = "bic")
fit_event <- fetwfeWithSimulatedData(
  sim,
  lambda_selection = "bic",
  fusion_structure = "event_study"
)

# The positions we wrote the truth to are exactly the ones fetwfe() reports as
# the treatment effects, so we can compare estimates to the truth cell by cell.
stopifnot(identical(treat_inds, fit_cohort$treat_inds))
```

How well does each fit recover the true per-cell effects? Compare the estimated
treatment-effect coefficients to the known truth (mean squared error over all
`r length(true_te)` cells):

```{r}
mse <- function(fit) mean((fit$beta_hat[treat_inds] - true_te)^2)
errs <- c(cohort = mse(fit_cohort), event_study = mse(fit_event))
round(errs, 3)
```

```{r, echo = FALSE}
pct <- round(100 * (1 - errs[["event_study"]] / errs[["cohort"]]))
```

On this panel the event-study penalty cuts the per-cell recovery error by about
`r pct`%. The advantage is exactly what we should expect: the truth *is* an
event-study structure, so pooling effects across cohorts at each event time pools
the right things, while the default penalty spends effort smoothing across
cohorts and calendar time that the truth does not actually share.

The same story is visible at the reporting level. Below we aggregate **both**
fits to one effect per event time with `eventStudy()` (the reporting family ---
applied here on top of both penalties) and overlay the true profile:

```{r, eval = requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("dplyr", quietly = TRUE), fig.width = 7, fig.height = 4.5}
library(ggplot2)
library(dplyr)

profiles <- dplyr::bind_rows(
  as.data.frame(eventStudy(fit_cohort)) |>
    dplyr::mutate(structure = "cohort (default)"),
  as.data.frame(eventStudy(fit_event)) |>
    dplyr::mutate(structure = "event_study")
)
# True profile per event time: the mean of the natively generated tau at each
# event time (the truth is event-study-sparse, not a clean formula line).
truth <- aggregate(
  estimate ~ event_time,
  data = data.frame(event_time = event_time, estimate = true_te),
  FUN = mean
)

ggplot(profiles, aes(event_time, estimate, color = structure, fill = structure)) +
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high), alpha = 0.15, color = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.6) +
  geom_line(
    data = truth, aes(event_time, estimate),
    color = "black", linetype = "dashed", inherit.aes = FALSE
  ) +
  labs(
    title = "Recovering an event-time treatment-effect profile",
    subtitle = "dashed black = true event-time profile (mean generated tau by event time)",
    x = "event time (periods since treatment)",
    y = "treatment effect",
    color = "fusion_structure", fill = "fusion_structure"
  ) +
  theme_minimal()
```

The event-study-penalized profile tracks the dashed truth more closely; the
default profile lags, because its calendar-and-cohort smoothing pulls effects
toward neighbors the truth does not actually share. (The exact magnitude --- and,
on any single draw, even the direction --- depends on the simulated data: across
event-study-sparse draws at this configuration the event-study penalty wins on
roughly four out of five seeds, and it is the per-cell coefficient recovery
above, not the aggregated profile, that captures that most reliably. The seed
shown here is one representative such draw.)

# When to prefer each

- **Prefer `fusion_structure = "event_study"`** when you believe treatment
  effects are driven mainly by **time since treatment** --- a gradual onset, a
  build-up or decay of the effect over exposure --- and you want to pool strength
  across cohorts at each event time. This matches the default null of the applied
  event-study literature, so the estimator's "shrink toward an event-study
  structure" behavior is a familiar story to report.

- **Prefer `fusion_structure = "cohort"` (the default)** when treatment effects
  plausibly **differ by cohort** --- cohort-specific responsiveness, or
  calendar-time shocks that hit all cohorts in a given year --- and you want
  smoothness across neighboring cohorts and within-cohort time without
  privileging an event-time summary. It is the right agnostic choice when you do
  not have a strong prior that event time is the only thing that matters.

Either way, the penalty is data-driven and the paper's guarantees hold for both,
so the decision affects efficiency, not validity. If you are unsure, the default
is a safe starting point; switch to `"event_study"` when the dynamic, time-since-
treatment path is the object of interest and you expect it to be shared across
cohorts.

# References

- Faletto, G. (2025). *Fused Extended Two-Way Fixed Effects for
  Difference-in-Differences With Staggered Adoptions.* arXiv preprint.
  <https://arxiv.org/abs/2312.05985>
- Callaway, B., & Sant'Anna, P. H. C. (2021). Difference-in-Differences with
  multiple time periods. *Journal of Econometrics*, 225(2), 200--230.
  <https://doi.org/10.1016/j.jeconom.2020.12.001>
