---
title: "High-dimensional inference: debiasedATT() when p >= NT"
author: "Gregory Faletto"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    math_method: "mathml"
vignette: >
  %\VignetteIndexEntry{High-dimensional inference: debiasedATT() when p >= NT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(fetwfe)
```

This vignette covers **high-dimensional inference** in `fetwfe`: what changes when the design has as many or more columns than observations ($p \geq NT$), and how `debiasedATT()` restores a uniformly valid confidence interval for the overall ATT.

It is a companion to `vignette("inference_vignette")` (how the package's standard errors are computed in the low-dimensional regime) and `vignette("simultaneous_cis_vignette")` (family-wise simultaneous bands).

# 1. High-dimensional inference ($p \geq NT$): `debiasedATT()`

The standard errors and simultaneous bands documented in the companion vignettes assume a **low-dimensional** design ($p < NT$), where the analytic Gram inverse exists and the fit's standard errors are the asymptotically valid ones of `vignette("inference_vignette")`. When the design is **high-dimensional** --- the number of design columns $p$ (cohort and time fixed effects, covariates, and all their treatment interactions) meets or exceeds the number of observations $NT$ --- two things change:

1. The fused fit is a *regularized* (bridge) estimate, and its pointwise standard errors (and the simultaneous bands of `vignette("simultaneous_cis_vignette")`) are **post-selection**: valid conditional on the selected support, not *uniformly* valid. They can under-cover, because the bridge's regularization biases the point estimate (it shrinks and prunes effects) and the post-selection standard error does not account for the selection.

2. `debiasedATT()` returns a **uniformly valid** confidence interval for the overall ATT by *desparsifying* the fit (the nodewise relaxed-inverse construction of the high-dimensional FETWFE theory in Faletto 2025), which removes the regularization bias. This is "one estimator, two regimes": the call and the returned object are identical to the $p < NT$ case; only the internal nuisance changes.

The $p \geq NT$ regime arises with many covariates, short panels, or many cohorts. `simulateData()` can generate such panels.

## A worked example

```{r highdim-setup}
# d = 12 covariates, N = 40 units, T = 5 periods, G = 3 cohorts.
coefs_hd <- genCoefs(G = 3, T = 5, d = 12, density = 0.2, eff_size = 2, seed = 2)
sim_hd <- simulateData(
  coefs_hd, N = 40, sig_eps_sq = 1, sig_eps_c_sq = 0.5, seed = 2
)
c(p = sim_hd$p, NT = sim_hd$N * sim_hd$T) # p >= NT: high-dimensional
```

```{r highdim-fit}
fit_hd <- fetwfeWithSimulatedData(sim_hd, q = 0.5)
# The fused (post-selection, pointwise) overall ATT:
round(c(att = fit_hd$att_hat, se = fit_hd$att_se), 3)
```

That fused estimate is post-selection. For uniformly valid inference, pass the fit to `debiasedATT()`, selecting the nodewise penalty by cross-validation (`lambda_c = "cv"`) --- the configuration whose overall-ATT coverage was validated at the anchor below. (The default fixed `lambda_c = 1.0` is a starting point, not the validated penalty.)

```{r highdim-debiased}
db <- debiasedATT(fit_hd, lambda_c = "cv")
db
```

The print surfaces the high-dimensional diagnostics worth inspecting: the resolved nodewise penalty (`lambda_c` / `lambda_node`) and, per debiasing direction, whether it **converged** and met its **KKT feasibility** certificate. A direction that fails to converge or stay feasible flags a potentially unreliable standard error --- raise `riesz_max_iter` and/or `lambda_c` and re-inspect.

## Why debiased, and how well it covers

In any single draw the fused and debiased point estimates can be close, so the value of `debiasedATT()` is not visible from one fit --- it is a **repeated-sampling** guarantee. A coverage study at a $p > NT$ anchor ([issue #88](https://github.com/gregfaletto/fetwfe/issues/88)) found:

| Overall-ATT interval | Coverage (nominal 95%) |
|---|---|
| Debiased (`debiasedATT()` with `lambda_c = "cv"`) | **≈ 93%** |
| Fused plug-in (pointwise)               | ≈ 77%                  |

The fused interval under-covers because of the post-selection non-uniformity; the debiased interval is near-nominal. This is validated at the studied anchor (with the feasibility-appropriate penalty); treat the $p \geq NT$ path as **experimental** outside that regime --- which is why the diagnostics above matter.

## An asymptotic alternative: the wild bootstrap

`method = "bootstrap"` offers an alternative, asymptotically-valid reference distribution for the overall-ATT interval: a studentized score / influence-function **wild cluster bootstrap** that re-signs the per-unit influence summands (no refit per replicate), leaving the point estimate and the reported `se` unchanged and changing only the critical value. It is **floored at the Gaussian quantile**, so the interval is never narrower than the analytic Wald interval.

**It is not a few-clusters remedy.** The analytic sandwich SE is downward-biased with few clusters, but this bootstrap does not fix that: as an *unrestricted, no-refit* score bootstrap it does not reproduce the tail inflation the *restricted* wild-cluster bootstrap-t uses for few-cluster coverage, and under heterogeneous cluster influence its critical value (before the floor) actually falls *below* the Gaussian --- so the floor reduces it to the analytic interval in exactly the small-`N` regime. Its only effect is to *widen* the interval when cluster influence is near-homogeneous. For genuine few-clusters inference a restricted bootstrap-t or a CR2-type analytic correction is required (future work).

```{r highdim-wild-boot, eval = FALSE}
# `fit` must be a single-sample fetwfe() fit -- not an `indep_counts` fit, which
# the worked example above (via fetwfeWithSimulatedData) happens to be.
debiasedATT(fit, method = "bootstrap", multiplier = "webb", seed = 1)
```

The `multiplier` (`"webb"`, the default; `"rademacher"`; or `"mammen"`) only affects the near-homogeneous case where the bootstrap widens above the floor --- in the small-`N` / heterogeneous regime the interval is the analytic one regardless of the weight. The bootstrap is defined only for **single-sample** fits (real observational panels are single-sample), not `indep_counts` two-sample fits.

## Simultaneous bands at $p \geq NT$

`simultaneousCIs()` also switches to the desparsified construction in the high-dimensional regime (via `method = "bootstrap"`; the analytic Gram-inverse band does not exist when $p \geq NT$):

```{r highdim-bands}
sc_hd <- simultaneousCIs(
  fit_hd,
  family = "cohort",
  method = "bootstrap",
  B = 500,
  seed = 1
)
sc_hd$regime
```

One scope caveat: #88 validated the **scalar overall-ATT** (`debiasedATT()`); the **family-wise band** coverage at $p \geq NT$ is not itself simulation-validated, so the band is the more experimental of the two.

## Controlled high-dimensional truth

The default `density` placement spreads signal across all coordinates, which when $p$ greatly exceeds the number of treatment-effect parameters rarely lands signal on the treatment-effect block. For a controlled, sparse, non-degenerate high-dimensional data-generating process, `genCoefs()` offers a **targeted-sparsity** mode (`n_signal_cohorts` or `treat_base_levels`) that places a heterogeneous treatment-effect signal directly on the per-cohort base levels --- the DGP the #88 coverage study uses. See `?genCoefs`.

# References

Faletto, G. (2025). "Fused Extended Two-Way Fixed Effects for Difference-in-Differences with Staggered Adoptions." *arXiv preprint* [arXiv:2312.05985](https://arxiv.org/abs/2312.05985).
