---
title: "Simulation Vignette for FETWFE: From Coefficients to True Treatment Effects"
author: "Gregory Faletto"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    output_file: "FETWFE_Simulation_Vignette.html"
vignette: >
  %\VignetteIndexEntry{FETWFE Simulation Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(echo = TRUE)

```

```{r setup}
# Load necessary libraries
library(dplyr)
library(fetwfe)
```

# Introduction

This vignette demonstrates how to perform simulation studies using the simulation functions provided in the `{fetwfe}` package. In particular, we will show how to:
  
- Generate a coefficient object with `genCoefs()`,
- Simulate a panel dataset with `simulateData()`,
- Run the FETWFE estimator on the simulated data using `fetwfeWithSimulatedData()`, and
- Extract the true treatment effects with `getTes()`.

We also illustrate how to use the native pipe operator (`|>`) to streamline the workflow.

# Simulation Workflow Using Piping

Below we create a complete simulation pipeline. First, we generate a coefficient object that contains the parameters for our simulation. Then, we simulate the panel data, and finally, we run the estimator on the simulated data. Separately, we also extract the true treatment effects for comparison.

## Step 1: Generate Simulation Coefficients

The `genCoefs()` function creates an object of class `"FETWFE_coefs"` which includes a coefficient vector along with the simulation parameters. For this example, we set:
  
- **R** (number of treated cohorts) to 3,
- **T** (number of time periods) to 10,
- **d** (number of covariates) to 2,
- **density** to 0.2 (controls sparsity), and
- **eff_size** to 1.5.

```{r}
# Generate the coefficient object for simulation
sim_coefs <- genCoefs(R = 5, T = 30, d = 12, density = 0.1, eff_size = 2, seed = 101)
```

## Step 2: Simulate Panel Data

Next, we simulate the data using the generated coefficient object with the `simulateData()` function. Here we specify:

- **N** (number of units) as 50,
- **sig_eps_sq** (observation-level noise variance) as 1,
- **sig_eps_c_sq** (unit-level noise variance) as 0.5, and
- use the default `"gaussian"` distribution for the covariates.

```{r}
# Simulate panel data based on the coefficients
sim_data <- simulateData(sim_coefs, N = 120, sig_eps_sq = 5, sig_eps_c_sq = 5, distribution = "gaussian")
```

## Step 3: Run the FETWFE Estimator on Simulated Data

We then run the estimator on the simulated data using `fetwfeWithSimulatedData()`. Using piping, the output from the previous step is seamlessly passed to the estimator.

```{r}
# Run the estimator using piping
result <- sim_data |>
  fetwfeWithSimulatedData()

# Display the estimated overall average treatment effect
cat("Estimated Overall ATT from Simulated Data:", result$att_hat, "\n")

# Calculate a 95% confidence interval
low_att <- result$att_hat - qnorm(1 - 0.05 / 2) * result$att_se
high_att <- result$att_hat + qnorm(1 - 0.05 / 2) * result$att_se

cat("Estimated Overall ATT from Simulated Data:")
cat(c(low_att, high_att))

# Print the estimated cohort-specific treatment effects
cat("Estimated cohort-specific treatment effects:")
result$catt_df

```

## Step 4: Extract True Treatment Effects

For comparison, we can compute the true treatment effects using the original coefficient object. The `getTes()` function extracts both the overall average treatment effect and the cohort-specific effects.

```{r}
# Extract the true treatment effects using piping
true_tes <- sim_coefs |> getTes()

# Print the true overall treatment effect
cat("True Overall ATT:", true_tes$att_true, "\n")

# Print the cohort-specific treatment effects
print(true_tes$actual_cohort_tes)
```

## Combining the Workflow in One Pipeline

Alternatively, you can chain the simulation functions together with the pipe operator. For example, the following code generates the coefficients, simulates the data, and runs the estimator all in one pipeline:

```{r}
# A fully piped workflow from coefficient generation to estimation
result_piped <- genCoefs(R = 5, T = 30, d = 12, density = 0.1, eff_size = 2, seed = 2025) |>
  simulateData(N = 120, sig_eps_sq = 5, sig_eps_c_sq = 5) |>
  fetwfeWithSimulatedData()

cat("Estimated Overall ATT from piped workflow:", result_piped$att_hat, "\n")
```

# Conclusion

In this vignette, we demonstrated how to use the simulation functions in the `{fetwfe}` package. By using piping, you can create a smooth workflow:
  
1. **Generate coefficients** with `genCoefs()`,
2. **Simulate data** using `simulateData()`,
3. **Run the FETWFE estimator** via `fetwfeWithSimulatedData()`, and
4. **Extract the true treatment effects** with `getTes()`.

This pipeline makes it easy to iterate on simulation studies and assess the performance of the estimator under different conditions. Feel free to modify the parameters and experiment with different settings!

For additional details or help, please refer to the package documentation or contact the author.
