---
title: "fetwfe: A Package for Fused Extended Two-Way Fixed Effects"
author: "Gregory Faletto"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    output_file: "FETWFE_vignette.html"
vignette: >
  %\VignetteIndexEntry{fetwfe: A Package for Fused Extended Two-Way Fixed Effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{dplyr}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Introduction

If you understand the basic idea of what difference-in-differences with staggered adoptions is, all you need to know about fused extended two-way fixed effects (FETWFE) to get started using the `{fetwfe}` package is this: given an appropriately formatted panel data set, `fetwfe()` will give you an estimate of the overall average treatment effect on the treated units, the average treatment effect within each cohort, and standard errors for each of these estimates.

Feel free to skip to the "Package Usage" section if you want to jump right in to using the package. In the next "Background" subsection, you can read a little more background information on the methodology if you'd like.

## Background

This vignette is written under the assumption that you're at least vaguely familiar with developments in [difference-in-differences](https://en.wikipedia.org/wiki/Difference_in_differences) with [staggered adoptions](https://bookdown.org/mike/data_analysis/difference-in-differences.html#staggered-dif-n-dif) since about 2018. Just to make sure we're on the same page, the brief recap is:

- Historically, under staggered adoptions researchers used the [standard two-way fixed effects estimator](https://bookdown.org/mike/data_analysis/two-way-fixed-effects.html) and interpreted the coefficient on the treatment dummy as an average treatment effect on the treated units.
- In the late 2010s, econometricians formally checked what this estimator was doing and found that in fact, this coefficient was not any kind of reasonable average treatment effect estimator.
- Since then, a number of new difference-in-differences estimators that are unbiased under staggered adoptions have been developed.

The estimator in this package, fused extended two-way fixed effects (FETWFE), is one of those unbiased estimators. Of course, I made this estimator because I think FETWFE brings something to the table that the others don't. Here's a brief summary on that:

One issue with these estimators has been that they've worked so hard to be unbiased that they are **inefficient** (in the language of econometrics), or **high-variance** (in the language of machine learning). These estimators add extra parameters in order to remove bias, but estimating extra parameters means you have less data per parameter and your estimates are noisier.

In machine learning, creating a more flexible estimator with lots of parameters and then finding that it is too high variance (that is, it *overfits*) is a familiar issue. The most common solution has been **regularization**.

You could just add $\ell_2$ or $\ell_1$ regularization to a difference-in-differences regression estimator and probably see an improvement in your efficiency, but FETWFE does something more sophisticated than that. (Plus, that approach wouldn't allow you to get valid standard errors for your treatment effect estimates, but FETWFE does.)

That's all the description I'll give you in this vignette. You can learn all of the details in the paper on arXiv:

> **[Fused Extended Two-Way Fixed Effects for Difference-in-Differences With Staggered Adoptions](https://arxiv.org/abs/2312.05985)**

If you want to learn a little more before you dive into the full paper, here are some other resources with descriptions of the methodology that provide a little more detail than this vignette:

- My [blog post](https://gregoryfaletto.com/2023/12/13/new-paper-fused-extended-two-way-fixed-effects-for-difference-in-differences-with-staggered-adoptions/) announcing the paper.
- [Some slides](https://gregoryfaletto.com/2024/02/11/presentation-on-fused-extended-two-way-fixed-effects/) I made for a presentation on FETWFE.
- Another [blog post](https://gregoryfaletto.com/2025/01/03/new-r-fetwfe-package-implementing-fused-extended-two-way-fixed-effects/) focused on what this package is doing under the hood.

But the headline summary of what fused extended two-way fixed effects brings to the table in a crowded field of estimators is: **fused extended two-way fixed effects is not only unbiased, it also uses machine learning to maximize efficiency (minimize variance)**. Further, unlike many machine learning estimators, **fused extended two-way fixed effects gives you valid standard errors for the treatment effect estimates.**

# Package Usage

The package provides a single exported function, `fetwfe()`, which implements the FETWFE estimator. Its primary arguments include:

- **`pdata`**: A data frame in panel (long) format.
- **`time_var`**: A character string specifying the name of the time period variable.
- **`unit_var`**: A character string specifying the unit (e.g. state, firm) variable.
- **`treatment`**: A character string specifying the treatment indicator variable (which must be an absorbing binary indicator).
- **`covs`**: A character vector of covariate names (typically time-invariant or the pre-treatment values).
- **`response`**: A character string specifying the response (outcome) variable.
- **Additional tuning parameters:** such as the tuning parameter for the bridge penalty (controlled via argument `q`) and options for verbosity, standard error calculation, and so on.

The function returns a list containing, for example, the estimated overall average treatment effect, cohort-specific treatment effects, standard errors (when available), and various diagnostic quantities.

You can get the full documentation details by using `?fetwfe` in R when you have the package loaded.

In the next sections, we'll walk through examples of how `fetwfe()` is used.

## Simulated Data Example

I'll start illustrating how to use `fetwfe()` by using a simulated data set. The example below simulates a balanced panel with 60 time periods, 30 individuals, and 5 waves of treatment. 

In the simulation, each individual is assigned a random cohort (which determines the timing of treatment) and three time-invariant covariates are generated. The response variable is constructed so that, after treatment, its evolution depends on a treatment effect (which varies by cohort) and a linear trend, plus the covariates and some random noise.

Below is the complete code for simulating the data, converting it into the required pdata format, and running the `fetwfe()` function.

**I borrowed some of the below code from [Asjad Naqvi](https://asjadnaqvi.github.io/)'s helpful [website for DiD estimators](https://asjadnaqvi.github.io/DiD/docs/code_r). Thanks for sharing the code publicly!**

```{r}
# Set seed for reproducibility
set.seed(123456L)

# 60 time periods, 30 individuals, and 5 waves of treatment
tmax = 60; imax = 30; nlvls = 5

dat = 
  expand.grid(time = 1:tmax, id = 1:imax) |>
  within({
    
    # Generate time-invariant covariates
    cov1 = rep(runif(imax, 0, 1), each = tmax)  # Random uniform values (0, 1) per individual
    cov2 = rep(sample(1:5, imax, replace = TRUE), each = tmax)  # Random categorical values (1-5) per individual
    cov3 = rep(rnorm(imax, mean = 0, sd = 1), each = tmax)  # Random Gaussian values (mean=0, sd=1) per individual
    
    # Initialize columns
    cohort      = NA
    effect      = NA
    first_treat = NA
    
    for (chrt in 1:imax) {
      cohort = ifelse(id==chrt, sample.int(nlvls, 1), cohort)
    }
    
    for (lvls in 1:nlvls) {
      effect      = ifelse(cohort==lvls, sample(2:10, 1), effect)
      first_treat = ifelse(cohort==lvls, sample(1:(tmax+20), 1), first_treat)
    }
    
    first_treat = ifelse(first_treat>tmax, Inf, first_treat)
    treat       = time >= first_treat
    rel_time    = time - first_treat
    y           = id + time + ifelse(treat, effect*rel_time, 0) + cov1 + cov2 + cov3 + rnorm(imax*tmax)
    
    rm(chrt, lvls, cohort, effect)
  })

head(dat)
```

The simulated data (`dat`) now has columns for time, id, covariates (`cov1`, `cov2`, `cov3`), a treatment indicator (`treat`), and a response variable (`y`). Next, we convert this data into the panel data format required by `fetwfe()`.

```{r}
library(dplyr)

# Specify column names for the pdata format
time_var <- "time"       # Column for the time period
unit_var <- "unit"       # Column for the unit identifier
treatment <- "treated"   # Column for the treatment dummy indicator
covs <- c("cov1", "cov2", "cov3")  # Columns for covariates
response <- "response"   # Column for the response variable

# Convert the dataset
pdata <- dat |>
  mutate(
    # Rename id to unit and convert to character
    {{ unit_var }} := as.character(id),
    # Ensure treatment dummy is 0/1
    {{ treatment }} := as.integer(treat),
    # Rename y to response
    {{ response }} := y
  ) |>
  select(
    {{ time_var }}, {{ unit_var }}, {{ treatment }}, all_of(covs), {{ response }}
  ) 

# Preview the resulting pdata dataframe
head(pdata)
```

Now that `pdata` is properly formatted, we run the FETWFE estimator on the simulated data.

```{r}

library(fetwfe)

# Run the FETWFE estimator on the simulated data
result <- fetwfe(
  pdata = pdata,              # The panel dataset
  time_var = "time",          # The time variable
  unit_var = "unit",          # The unit identifier
  treatment = "treated",      # The treatment dummy indicator
  covs = c("cov1", "cov2", "cov3"),  # Covariates
  response = "response"      # The response variable
)

# Display the overall average treatment effect estimate
cat("Estimated Overall ATT:", result$att_hat, "\n")
```

When you run this code, the function internally performs all the necessary data preparation, applies the fusion penalty via a bridge regression (using the `grpreg` package), and returns a list with overall and cohort-specific treatment effect estimates, standard errors (if available), and additional diagnostics.

## A "Real Data" Example

Next I illustrate FETWFE in an empirical context. I'll use data from Stevenson and Wolfers (2006), via the `divorce` data set from the `bacondecomp` package, on the impact of no-fault divorce laws on women's suicide rates. (See also Goodman-Bacon (2021) for an alternative analysis.) The below is identical to the data application from [my paper](https://arxiv.org/abs/2312.05985).

In this application, the panel data consist of state-level observations over 33 years. After removing states that received treatment in the first period, we are left with 42 states, of which 5 are never treated and 12 cohorts adopt treatment at various times.

Time-varying covariates (such as the state homicide rate, logged personal income, and welfare participation) are included as controls (using the pre-treatment values). In this example, I use FETWFE to estimate the marginal average treatment effect of no-fault divorce laws. Note: the below code takes about 15 seconds to run on my laptop.

```{r}
library(bacondecomp)  # for the example data

# Load the example data
data(divorce)

set.seed(23451)

# Suppose we wish to estimate the effect of a policy (here represented by the variable "changed")
# on the response "suiciderate_elast_jag" using covariates "murderrate", "lnpersinc", and "afdcrolls".
# Here
# - 'year' is the time period variable (as an integer),
# - 'st' is the unit identifier,
# - 'changed' is the treatment indicator (with 0 = untreated, 1 = treated),
# 
# The `fetwfe()` function will automatically take care of removing units that were treated in the
# first time period.

# Call the estimator
res <- fetwfe(
    pdata=divorce[divorce$sex == 2, ],
    time_var="year",
    unit_var="st",
    treatment="changed",
    covs=c("murderrate", "lnpersinc", "afdcrolls"),
    response="suiciderate_elast_jag"
    )

# Average treatment effect on the treated units (in percentage point
# units)
100 * res$att_hat

# Conservative 95% confidence interval for ATT (in percentage point units)

low_att <- 100 * (res$att_hat - qnorm(1 - 0.05 / 2) * res$att_se)
high_att <- 100 * (res$att_hat + qnorm(1 - 0.05 / 2) * res$att_se)

c(low_att, high_att)

# Cohort average treatment effects and confidence intervals (in percentage
# point units)

catt_df_pct <- res$catt_df
catt_df_pct[["Estimated TE"]] <- 100 * catt_df_pct[["Estimated TE"]]
catt_df_pct[["SE"]] <- 100 * catt_df_pct[["SE"]]
catt_df_pct[["ConfIntLow"]] <- 100 * catt_df_pct[["ConfIntLow"]]
catt_df_pct[["ConfIntHigh"]] <- 100 * catt_df_pct[["ConfIntHigh"]]

catt_df_pct
```

For the data application, FETWFE yielded an overall ATT of approximately –3.76% change in the female suicide rate, similar to other estimates in the literature. In addition, the output table (stored in `result_emp$catt_df`) displays the cohort-specific estimates. (Note that standard errors for the individual cohort estimates are less reliable when the number of units per cohort is small.)

# Conclusion

This should be enough to get you started using `fetwfe()` on your own data. Please feel free to [reach out](https://gregoryfaletto.com/about/) if you have any questions or feedback or run into any issues using the package. You can also [create an issue](https://github.com/gregfaletto/fetwfePackage/issues) if you think there’s a bug in the package or you’d like to request a feature. **Thanks so much for checking out the package!**

# References

- Faletto, G. (2024). Fused Extended Two-Way Fixed Effects for Difference-in-Differences with Staggered Adoptions. [arXiv preprint arXiv:2312.05985](https://arxiv.org/abs/2312.05985).
- Goodman-Bacon, A. (2021). Difference-in-Differences with variation in treatment timing. *Journal of Econometrics*.
- Flack, E., & Jee, E. (2020). bacondecomp: Goodman-Bacon Decomposition. R package version 0.1.1. [https://CRAN.R-project.org/package=bacondecomp](https://CRAN.R-project.org/package=bacondecomp).
- Stevenson, B., & Wolfers, J. (2006). Bargaining in the shadow of the law: Divorce laws and family distress. *The Quarterly Journal of Economics*, 121(1), 267-288.

