---
title: "fetwfe: A Package for Fused Extended Two-Way Fixed Effects"
author: "Gregory Faletto"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    output_file: "FETWFE_vignette.html"
vignette: >
  %\VignetteIndexEntry{fetwfe: A Package for Fused Extended Two-Way Fixed Effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{dplyr}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Introduction

This vignette is written under the assumption that you're vaguely familiar with developments in [difference-in-differences](https://en.wikipedia.org/wiki/Difference_in_differences) with [staggered adoptions](https://bookdown.org/mike/data_analysis/difference-in-differences.html#staggered-dif-n-dif) since about 2018. Just to make sure we're on the same page, the brief recap is:
* Historically, under staggered adoptions researchers used the [standard two-way fixed effects estimator](https://bookdown.org/mike/data_analysis/two-way-fixed-effects.html) and interpreted the coefficient on the treatment dummy as an average treatment effect on the treated units.
* In the late 2010s, econometricians formally checked what this estimator was doing and found that in fact, this coefficient was not any kind of reasonable average treatment effect estimator.
* Since then, a number of new difference-in-differences estimators that are unbiased under staggered adoptions have been developed.

The estimator in this package, fused extended two-way fixed effects (FETWFE), is one of those unbiased estimators. One issue with these estimators has been that they've worked so hard to be unbiased that they are inefficient (in the language of econometrics), or high-variance (in the language of machine learning). In particular, these estimators add extra parameters in order to remove bias, but estimating extra parameters means you have less data per parameter and your estimates are noisier.

In the machine learning literature, creating a more flexible estimator with lots of parameters and then finding that it is too high variance (that is, it *overfits*) is a familiar issue. The most common solution has been regularization.

So the headline summary of what fused extended two-way fixed effects brings to the table in a crowded field of estimators is: **fused extended two-way fixed effects is not only unbiased, it also uses machine learning to maximize efficiency (minimize variance)**. Further, unlike many machine learning estimators, **fused extended two-way fixed effects still gives you standard errors for the treatment effect estimates.**

That's all the description I'll give you in this vignette. You can learn all of the details in the paper on arXiv:

> **[Fused Extended Two-Way Fixed Effects for Difference-in-Differences With Staggered Adoptions](https://arxiv.org/abs/2312.05985)**

If you want to learn a little more before you dive into the full paper, here are some other resources with descriptions of the methodology that provide a little more detail than this vignette:
* My [blog post](https://gregoryfaletto.com/2023/12/13/new-paper-fused-extended-two-way-fixed-effects-for-difference-in-differences-with-staggered-adoptions/) announcing the paper.
* [Some slides](https://gregoryfaletto.com/2024/02/11/presentation-on-fused-extended-two-way-fixed-effects/) I made for a presentation on FETWFE.
* Another [blog post](https://gregoryfaletto.com/2025/01/03/new-r-fetwfe-package-implementing-fused-extended-two-way-fixed-effects/) focused on what this package is doing under the hood.

For now, all you need to know is that given an appropriately formatted panel data set, `fetwfe()` will give you an estimate of the overall average treatment effect on the treated units, the average treatment effect within each cohort, and standard errors for each of these estimates.

# Package Usage

The package provides a single exported function, `fetwfe()`, which implements the FETWFE estimator. Its primary arguments include:

- **`pdata`**: A data frame in panel (long) format.
- **`time_var`**: A character string specifying the name of the time period variable.
- **`unit_var`**: A character string specifying the unit (e.g. state, firm) variable.
- **`treatment`**: A character string specifying the treatment indicator variable (which must be an absorbing binary indicator).
- **`covs`**: A character vector of covariate names (typically time-invariant or the pre-treatment values).
- **`response`**: A character string specifying the response (outcome) variable.
- **Additional tuning parameters:** such as the tuning parameter for the bridge penalty (controlled via argument `q`) and options for verbosity, standard error calculation, and so on.

The function returns a list containing, for example, the estimated overall average treatment effect, cohort-specific treatment effects, standard errors (when available), and various diagnostic quantities.

## Simulated Data Example

In addition to real data applications, you can also experiment with FETWFE on simulated data. The example below simulates a balanced panel with 60 time periods, 30 individuals, and 5 waves of treatment. In the simulation, each individual is assigned a random cohort (which determines the timing of treatment) and three time-invariant covariates are generated. The response variable is constructed so that, after treatment, its evolution depends on a treatment effect (which varies by cohort) and a linear trend, plus the covariates and some random noise.

Below is the complete code for simulating the data, converting it into the required pdata format, and running the `fetwfe()` function.

```{r}
# Set seed for reproducibility
set.seed(123456L)

# 60 time periods, 30 individuals, and 5 waves of treatment
tmax = 60; imax = 30; nlvls = 5

dat = 
  expand.grid(time = 1:tmax, id = 1:imax) |>
  within({
    
    # Generate time-invariant covariates
    cov1 = rep(runif(imax, 0, 1), each = tmax)  # Random uniform values (0, 1) per individual
    cov2 = rep(sample(1:5, imax, replace = TRUE), each = tmax)  # Random categorical values (1-5) per individual
    cov3 = rep(rnorm(imax, mean = 0, sd = 1), each = tmax)  # Random Gaussian values (mean=0, sd=1) per individual
    
    # Initialize columns
    cohort      = NA
    effect      = NA
    first_treat = NA
    
    for (chrt in 1:imax) {
      cohort = ifelse(id==chrt, sample.int(nlvls, 1), cohort)
    }
    
    for (lvls in 1:nlvls) {
      effect      = ifelse(cohort==lvls, sample(2:10, 1), effect)
      first_treat = ifelse(cohort==lvls, sample(1:(tmax+20), 1), first_treat)
    }
    
    first_treat = ifelse(first_treat>tmax, Inf, first_treat)
    treat       = time >= first_treat
    rel_time    = time - first_treat
    y           = id + time + ifelse(treat, effect*rel_time, 0) + cov1 + cov2 + cov3 + rnorm(imax*tmax)
    
    rm(chrt, lvls, cohort, effect)
  })

head(dat)
```

The simulated data (`dat`) now has columns for time, id, covariates (`cov1`, `cov2`, `cov3`), a treatment indicator (`treat`), and a response variable (`y`). Next, we convert this data into the panel data format required by `fetwfe()`.

```{r}
library(dplyr)

# Specify column names for the pdata format
time_var <- "time"       # Column for the time period
unit_var <- "unit"       # Column for the unit identifier
treatment <- "treated"   # Column for the treatment dummy indicator
covs <- c("cov1", "cov2", "cov3")  # Columns for covariates
response <- "response"   # Column for the response variable

# Convert the dataset
pdata <- dat |>
  mutate(
    # Rename id to unit and convert to character
    {{ unit_var }} := as.character(id),
    # Ensure treatment dummy is 0/1
    {{ treatment }} := as.integer(treat),
    # Rename y to response
    {{ response }} := y
  ) |>
  select(
    {{ time_var }}, {{ unit_var }}, {{ treatment }}, all_of(covs), {{ response }}
  ) 

# Preview the resulting pdata dataframe
head(pdata)
```

Now that `pdata` is properly formatted, we run the FETWFE estimator on the simulated data.

```{r}

library(fetwfe)

# Run the FETWFE estimator on the simulated data
result <- fetwfe(
  pdata = pdata,              # The panel dataset
  time_var = "time",          # The time variable
  unit_var = "unit",          # The unit identifier
  treatment = "treated",      # The treatment dummy indicator
  covs = c("cov1", "cov2", "cov3"),  # Covariates
  response = "response",      # The response variable
  q = 0.5                    # The L_q penalty for fusion regularization
)

# Display the overall average treatment effect estimate
cat("Estimated Overall ATT:", result$att_hat, "\n")
```

When you run this code, the function internally performs all the necessary data preparation, applies the fusion penalty via a bridge regression (using the `grpreg` package), and returns a list with overall and cohort-specific treatment effect estimates, standard errors (if available), and additional diagnostics.

This example shows how you can simulate panel data with treatment heterogeneity, convert it to the required format, and apply the FETWFE estimator to recover treatment effect estimates.

## A Basic Example

Below is an example of how to use `fetwfe()` on a panel data set. (In this example I use the `divorce` data from the `bacondecomp` package; you may substitute any appropriate panel data.)

```{r}
# Load the necessary packages
library(fetwfe)
library(bacondecomp)  # for the example data

# Load the example data
data(divorce)

# Suppose we wish to estimate the effect of a policy (here represented by the variable "changed")
# on the response "suiciderate_elast_jag" using covariates "murderrate", "lnpersinc", and "afdcrolls".
# Assume that:
# - 'year' is the time period variable (as an integer),
# - 'st' is the unit identifier,
# - 'changed' is the treatment indicator (with 0 = untreated, 1 = treated),
# - We only use observations that are not treated in the first period.

# Call the estimator (note: adjust variable names as needed)
result <- fetwfe(
    pdata=divorce[divorce$sex == 2, ],
    time_var="year",
    unit_var="st",
    treatment="changed",
    covs=c("murderrate", "lnpersinc", "afdcrolls"),
    response="suiciderate_elast_jag",
    q=0.5 # fusion penalty parameter (q in (0,2])
    )

# Display the overall average treatment effect
cat("Overall ATT estimate:", result$att_hat, "\n")

# If available, display the conservative 95% confidence interval
if (!is.na(result$att_se)) {
  se <- result$att_se
  lower <- result$att_hat - qnorm(0.975) * se
  upper <- result$att_hat + qnorm(0.975) * se
  cat("95% Confidence Interval: (", lower, ", ", upper, ")\n")
}

# Display the cohort-specific treatment effects as a data frame
print(result$catt_df)
```

When you run this code, the function internally performs all necessary data transformations, applies the fusion penalty via a bridge regression (using the `grpreg` package), and returns the estimated treatment effects along with diagnostic information.

# Empirical Application

To illustrate FETWFE in an empirical context, we now reproduce an analysis using data from Stevenson and Wolfers (2006) on the impact of no-fault divorce laws on female suicide rates. (See also Goodman-Bacon (2021) for an alternative analysis.) In our application, the panel data consist of state-level observations over 33 years. After removing states that received treatment in the first period, we are left with 42 states, of which 5 are never treated and 12 cohorts adopt treatment at various times.

Time-varying covariates (such as the state homicide rate, logged personal income, and welfare participation) are included as controls (using the pre-treatment values). In our example, we use FETWFE with \(q = 0.5\) to estimate the marginal average treatment effects. For instance:

```{r}
# Load the divorce data (from bacondecomp package)
data(divorce)

# (For illustration: filter out units treated in the first period if necessary)
# In our empirical application, suppose we remove states treated at t = 1.
# This step is handled internally by the function, but you can also pre-filter your data.

# Call the FETWFE estimator with appropriate parameters.
# (Assume that the following variable names correspond to the variables described in the paper.)
result_emp <- fetwfe(
  pdata=divorce[divorce$sex == 2, ],
  time_var="year",
  unit_var="st",
  treatment="changed",
  covs=c("murderrate", "lnpersinc", "afdcrolls"),
  response="suiciderate_elast_jag",
  q=0.5
)

# Print the overall average treatment effect (ATT)
cat("Estimated overall ATT (percent change):", 100 * result_emp$att_hat, "\n")

# Print the cohort-specific treatment effect estimates
print(result_emp$catt_df)
```

For the data application, our estimates yielded an overall ATT of approximately –3.76% change in the female suicide rate, similar to other estimates in the literature. In addition, the output table (stored in `result_emp$catt_df`) displays the cohort-specific estimates. (Note that standard errors for the individual cohort estimates are less reliable when the number of units per cohort is small.)

Below is an excerpt of the table produced by our estimator:

| Cohort | Estimated TE |   SE   | ConfIntLow | ConfIntHigh |
|:------:|:------------:|:------:|:----------:|:-----------:|
| 1969   |  0%          |   —    |     —      |     —       |
| 1970   | -40.142%     |   —    |     —      |     —       |
| 1971   |  0%          |   —    |     —      |     —       |
| ⋮      |   ⋮          |   ⋮    |     ⋮      |     ⋮       |

*Table: Cohort average treatment effect estimates.*

# Conclusion

This vignette has introduced the FETWFE estimator—a fusion-penalized extended two-way fixed effects method—for estimating treatment effects in panel data with staggered adoptions. By automatically fusing together similar treatment effects across cohorts and time periods, FETWFE reduces model complexity while preserving desirable asymptotic properties. In our examples, we demonstrated how to use the single function `fetwfe()` (or `fewtfe()` as you may name it) to obtain both overall and cohort-specific treatment effect estimates.

I encourage users to explore further the detailed theoretical properties (consistency, oracle property, asymptotic normality) presented in the accompanying paper. Feedback and suggestions are welcome!

# References

- Borusyak, K., Jaravel, X., & Spiess, J. (2024). Revisiting difference-in-differences. *Journal of Econometrics*.
- Callaway, B., & Sant’Anna, P. H. (2021). Difference-in-Differences with multiple time periods. *Journal of Econometrics*.
- Goodman-Bacon, A. (2021). Difference-in-Differences with variation in treatment timing. *Journal of Econometrics*.
- Huang, J., Horowitz, J. L., & Wei, F. (2008). Asymptotic properties of bridge estimators in sparse high-dimensional regression models. *Annals of Statistics*.
- Pesaran, M. H. (2015). *Time Series and Panel Data Econometrics*. Oxford University Press.
- Sun, L., & Abraham, S. (2021). Estimating dynamic treatment effects in event studies with heterogeneous treatment effects. *Journal of Econometrics*.
- Tibshirani, R. (1996). Regression shrinkage and selection via the lasso. *Journal of the Royal Statistical Society: Series B*.
- Tibshirani, R., Saunders, M., Rosset, S., Zhu, J., & Knight, K. (2005). Sparsity and smoothness via the fused lasso. *Journal of the Royal Statistical Society: Series B*.
- Wooldridge, J. M. (2021). Two-way fixed effects estimators with heterogeneous treatment effects. *Journal of Econometrics*.

